Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions

### 2025-01-20 src/article.tex
Updates from Overleaf

This work inherits the same philosophy as the Tromba Moderna project. By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions. Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}. As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to adhere to longstanding harpsichord building traditions. However, this work complements and follows up on previous reflections regarding the \emph{sustainability of results} within the NIME community and identified as \emph{the O in NIME} \cite{Masu_NIME_2023}. The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.

Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through upcoming projects such as the ERC-funded Synergy grant Rem@ke \cite{remake1}.

@@

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\textbackslash FIGURE, but there was still a limit given the structural
and aesthatic requirements.

@@

Deviations from the design of the Trasuntino are the rectangular, rather than
logartithmic shape, removal of a bottom panel and ommision of a short octave,
instead spanning 4 octaves beginning on C.

The strings are Malcolm Rose yellow brass wire, which are secured to wrought
iron pins. The tension of each string was adjusted to provide resistance to
plucking close to that of a harpsichord. Three felt strips were woven between
the strings so as to cancel the vibration of the strings.

@@

\begin{itemize}
\item
  Non-invasive: The sensor system should not require major modifications
  to the interface on which it is installed.
\item
  Low Latency: Reading sensors and then triggering an output should have
  a low latency. Using {[}paper on latency{]} a target was set as X ms
  with an upper limit set to 25ms beyond which the sensors would be
  deemed unusable
\item
  Reliable: Data obtained should be reliable and accurate enough not to
  result in false positives or false negatives without extensive
  filtering
\item
  Scalable: The system should be scalable both economically and
  temporally. Given the open soucre nature of the porject, it could not
  be prohibitvely expensive to carry out. Also any system that would
  work on a single key would have to scale financially to potentially 50
  or more. The time required to install and calibrate the sensors would
  also have to scale.
\item
  Expandable: Any sensor system should be functionally expandable to
  allow for the usage of the data for control over other parameters.
\end{itemize}

>>>>

This work inherits the same philosophy as the Tromba Moderna project. By augmenting a replica of a historical harpsichord keyboard using minimally invasive electronics and controlling a MIDI-triggered sample library, the project intends to offer a tool enhancing the fruition of the Tagliavini collection whilst retaining a form of continuity with historical instrument-building traditions. Furthermore, the electronics design borrows ideas from a previous NIME by McPherson \cite{McPherson2013}. As such, the interface proposed here is not, strictly speaking, a \emph{new} musical interface, particularly in its haptic response designed to adhere to longstanding harpsichord building traditions. However, this work complements and follows up on previous reflections regarding the \emph{sustainability of results} within the NIME community and identified as \emph{the O in NIME} \cite{Masu_NIME_2023}. The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development.

Beyond enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anona{ERC-funded NEMUS} project, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1}.

@@

Accommodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the electronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\textbackslash FIGURE, but there was still a limit given the structural
and aesthetic requirements.

@@

Deviations from the design of the Trasuntino are the rectangular, rather than
'logarithmic' shape, removal of a bottom panel and omission of a short octave,
instead spanning 4 octaves beginning on C.

The strings are Malcolm Rose yellow brass wire, which are secured to wrought
iron pins. The tension of each string was adjusted to provide resistance to
plucking close to that of a harpsichord. Three felt strips were woven between
the strings in order to dampen the vibration of the strings.

@@

\begin{itemize}
\item
  Non-invasive: The sensor system should not require major modifications
  to the interface on which it is installed.
\item
  Low Latency: Reading sensors and then triggering an output should have
  a low latency. Following \cite{Jack2016} an latency was set as 10 ms
  with an upper limit set to 25ms beyond which the sensors would be
  deemed unusable
\item
  Reliable: Data obtained should be reliable and accurate enough not to
  result in false positives or false negatives without extensive
  filtering
\item
  Scalable: The system should be scalable both economically and
  temporally. Given the open source nature of the project, it could not
  be prohibitively expensive to carry out. Also any system that would
  work on a single key would have to scale financially to potentially 50
  or more. The time required to install and calibrate the sensors would
  also have to scale.
\item
  Expandable: Any sensor system should be functionally expandable to
  allow for the usage of the data for control over other parameters.
\end{itemize}
